---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---
# parameters  
```{r}
update_eye_data <- FALSE
```


# Data Preparation 
## load packages

```{r}
library(lubridate)
library(here)
library(tidyverse)
library(eyetrackingR)
library(janitor)
library(readxl)
library(rio)
library(zoo)
library(anytime)
library(tidylog)
```

## read eye data
```{r}
data <- readRDS(here("results", "processed_data", "cm_merged_eye_data.Rdata"))
```

## read participant info
```{r}
combined_info <- readRDS(here("results", "processed_data", "cm_participant_info.Rdata"))
```




## read information on images presented & AOIs
```{r}
# side of presentation for each object
images <- read.csv(here("data", "trial_info", "trial_info.csv"))
item_info <- read_excel(here("data", "trial_info", "item_info.xlsx"))
# target AOI
aoi_target <- read.csv(here("data", "trial_info", "aoi_target.csv"))
# distractor AOI
aoi_distractor <- read.csv(here("data", "trial_info", "aoi_distractor.csv"))
```



# work on participant info
## create language group variable
```{r}
combined_info <- combined_info %>%
  mutate(lang_group = case_when(
    percent_english >= 90 ~ "monolingual",
    percent_english >= 25 & percent_french >= 25 & percent_other < 10 ~ "bilingual",
    TRUE ~ "none"
  ))

# very stupid graph to check that assignment is correct
ggplot(combined_info, aes(
  x = percent_english,
  y = percent_french,
  colour = lang_group
)) +
  geom_point() +
  facet_grid(percent_other > 10 ~ .)

combined_info %>%
  group_by(lang_group) %>%
  count()
```

## participant exclusion



## premature birth
```{r}
combined_info <- combined_info %>%
  mutate(exclude_premature = case_when(
    gestation_weeks < 37 | gestation == "Prem" ~ 1,
    weight_in_grams < 2500 ~ 1,
    is.na(weight_in_grams) & is.na(gestation_weeks) ~ as.numeric(NA),
    TRUE ~ 0
  ))
# noticed during code review: there is one kid (P19) getting NA for exclude premature, but I checked their demographics form and the parent answered "no" for being born early 
# to fix this:
combined_info = combined_info %>%
  mutate(exclude_premature = ifelse(part_id == "P19", 0, exclude_premature))

```

## health 
```{r}

combined_info <- combined_info %>%
  mutate(exclude_health = NA)


combined_info %>%
  distinct(health_comments, .keep_all = T) %>%
  select(part_id, health_comments) %>%
  write_csv(here("results", "processed_data", "to_check", "cm_health_to_check.csv"))


health_comments_checked <- read_csv(here("results", "processed_data", "to_check", "cm_health_checked.csv"))

# merge back into combined_info and create exclude_health variabl
combined_info <- combined_info %>%
  left_join(health_comments_checked %>% select(-health_comments)) %>%
  mutate(exclude_health = ifelse(exclude_health_comments == FALSE | is.na(health_comments), 0, 1))
```

## language

```{r}
combined_info <- combined_info %>%
  mutate(exclude_language = case_when(
    lang_group == "none" ~ 1,
    # participants who did not hear English from birth OR French from birth (ie. sequential bilinguals)
    part_id %in% c("P22", "P68") ~ 1,
    !is.na(lang_group) ~ 0,
    TRUE ~ as.numeric(NA)
  ))
```





## age
```{r}
combined_info <- combined_info %>%
  mutate(
    exclude_age = ifelse(between(age_months, 23, 33), 0, 1)
  )
```


## overview of participant exclusions
```{r}
combined_info <- combined_info %>%
  mutate(exclude_summary = case_when(
    exclude_age == 1 ~ "1_age",
    exclude_premature == 1 ~ "2_preterm",
    exclude_health == 1 ~ "3_health",
    exclude_language == 1 ~ "4_language",
    TRUE ~ "keeper"
  ))

combined_info %>%
  group_by(exclude_summary) %>%
  count()
```
# keep only necessary variables for analysis
```{r}
combined_info <- combined_info %>%
  select(
    study, part_id, gender, percent_english, percent_french, percent_other, percent_dominant, percent_nondominant, dominant_language, exclude_summary, lang_group, age_days, age_months, child_ethnicity, years_education,
    language_mixing_score, cdi_forms_filled_visit, cdi_responding_caregiver, cdi_vocab_production_en, cdi_vocab_production_fr
  )
```


# eyetracking data prep

## create necessary variables 
### trial info from soundfile name etc. 
```{r}


data <- data %>% rename(
  gaze_x = gaze_point_x_adc_spx,
  gaze_y = gaze_point_y_adc_spx,
  order = studio_test_name,
  timestamp = recording_timestamp
) ## Renaming the column


data$trial_attg <- ifelse(startsWith(as.character(data$media_name), "Attentiongetter"),
  "att",
  "test"
)
data$trial_attg <- ifelse(data$media_name == "", "other", data$trial_attg) 


data$movie_name <- ifelse(data$trial_attg == "test",
  as.character(data$media_name),
  ""
)
data$movie_name <- gsub(".avi", "", data$movie_name)
# this throws error message because on attentiongetter trials it fails
data <- data %>% separate(movie_name,
  into = c(NA, "image", "audio", "order_trial"),
  sep = "_",
  remove = F
)
data$trial <- substr(data$order_trial, start = 3, stop = 5)
data$trial_num <- as.numeric(substr(data$order_trial, start = 4, stop = 5))
data$object_named <- substr(data$audio, 1, 3)
# create trial type variables
data$misp_cond <- substr(data$audio, 4, 4)
data$word_type <- substr(as.character(data$image), 1, 1)
data <- data %>%
  mutate(
    misp_cond = case_when(
      substr(audio, 4, 4) == "C" ~ "correct",
      substr(audio, 4, 4) == "M" ~ "mispronounced"
    ),
    word_type = case_when(
      substr(as.character(data$image), 1, 1) == "n" ~ "non-cognate",
      substr(as.character(data$image), 1, 1) == "c" ~ "cognate"
    ),
    trial_type = case_when(
      misp_cond == "correct" & word_type == "non-cognate" ~ "correctly pronounced non-cognate",
      misp_cond == "mispronounced" & word_type == "non-cognate" ~ "mispronounced non-cognate",
      misp_cond == "correct" & word_type == "cognate" ~ "correctly pronounced cognate",
      misp_cond == "mispronounced" & word_type == "cognate" ~ "mispronounced cognate"
    )
  )


data$item <- substr(as.character(data$image), 1, 2)
data$item_named <- paste0(data$item, data$object_named)
```




### add info on images presented & AOIS
```{r}
# change order labels from e1 to E1 for compability with eyetracking data
images <- images %>% mutate(order = toupper(order))

# add info on which image is being seen

data <- full_join(data, images)
```
### add info on items
```{r}
data <- data %>% left_join(item_info %>% select(
  item_named_full,
  misp_type,
  item_named
))
```


#### check that matching worked correctly
```{r}
data <- data %>% mutate(
  item_code = gsub("_", "", item_code),
  matching_item = ifelse(item_code == image, "ok", "check")
)
# check that when there is a trial going on (trial_attg =="test"), all are "ok"
data %>%
  group_by(trial_attg, matching_item) %>%
  count()
```


## dealing with track loss
```{r}

data <- data %>% mutate(
  track_loss_reason =
    case_when(
      between(as.numeric(as.character(validity_right)), 1, 4) & between(as.numeric(as.character(validity_left)), 1, 4) ~ "validity is bad", 
      is.na(gaze_x) | is.na(gaze_y) ~ "no info on coordinates",
      gaze_x < 0 | gaze_x > 1920 ~ "outside of box", 
      gaze_y < 0 | gaze_y > 1200 ~ "outside of box",
      between(gaze_x, 0, 1920) ~ "",
      between(gaze_y, 0, 1200) ~ "",
      TRUE ~ "not sure"
    ),
  track_loss = ifelse(track_loss_reason != "", TRUE, FALSE)
)
data %>% tabyl(track_loss_reason)
```


## add AOI info
```{r}

# is target L or R?
data <- data %>%
  mutate(target_location = case_when(
    object_named == image_left ~ "target_left",
    object_named == image_right ~ "target_right",
    TRUE ~ as.character(NA)
  ))

# add target aoi
data <- add_aoi(data,
  aoi_target,
  x_col = "gaze_x",
  y_col = "gaze_y",
  aoi_name = "target"
)
# add distractor aoi
data <- add_aoi(data,
  aoi_distractor,
  x_col = "gaze_x",
  y_col = "gaze_y",
  aoi_name = "distractor"
)
```


## check that data are on the right coordinate system
```{r}


## Finalizing data


# get rid of rows for attentiongetter &  where no video is presented
# but want to keep row for part_id P76 who doesn't have eyetracking data so we keep this participant in stats about excluded participants
data <- data %>% filter(
  trial_attg != "att" | part_id == "P76",
  !is.na(media_name) | part_id == "P76",
  media_name != "" | part_id == "P76"
)



data$gaze_y <- (data$gaze_y - 1200) * -1 # Put Gaze_y on a cartesian coordinate system instead of Tobii system

# the following code takes a bit to run
# check that there are no outliers left ( all gaze points are y:0-1200 and y 0-1900)
# with(data[data$track_loss==F,],     plot(gaze_x, gaze_y)) 
```



# combine eyetracking data with participant info
## test merging to find mismatches
```{r}
missing_from_info <- anti_join(
  data %>% distinct(part_id),
  combined_info
) %>% print()

# one participant (P76) does not have eyetracking data, but added that one as a row on data, so doesn't show up here
missing_from_eyetracker <- anti_join(
  combined_info,
  data %>% distinct(part_id)
) %>% print()


full_data <- left_join(data, combined_info)
```
# eyetracking data preprocessing
## eyetrackingR preparations
```{r}
# delete duplicated rows (artifact introduced by tobii eyetracker)
full_data_dup_removed <- full_data %>%
  arrange(part_id, trial_num, timestamp, studio_event) %>%
  distinct(part_id, trial_num, timestamp, .keep_all = T)

data_eye_r <- make_eyetrackingr_data(full_data_dup_removed,
  participant_column = "part_id",
  trial_column = "trial_num",
  time_column = "timestamp",
  trackloss_column = "track_loss",
  aoi_columns = c("target", "distractor"),
  treat_non_aoi_looks_as_missing = TRUE
)

# check that every participant & trial has moviestart & movieend message
full_data_dup_removed %>%
  group_by(part_id, media_name, studio_event) %>%
  count() %>%
  spread(studio_event, n)
```

## rezero data
```{r}
# create a trial-specific timestamp (instead of timestamp relative to start of study)
data_rezero <- subset_by_window(data_eye_r,
  window_start_msg = "MovieStart",
  window_end_msg = "MovieEnd",
  msg_col = "studio_event",
  rezero = T
)

data_rezero <- make_eyetrackingr_data(data_rezero,
  participant_column = "part_id",
  trial_column = "media_name",
  time_column = "timestamp",
  trackloss_column = "track_loss",
  aoi_columns = c("target", "distractor"),
  treat_non_aoi_looks_as_missing = TRUE
)

# add row for P76 back in so we can count this person in excluded participant summary in cm4
data_rezero <- data_rezero %>%
  full_join(data_eye_r %>% filter(part_id == "P76"))
```

# save data for further analysis
```{r}
saveRDS(data_rezero, here("results", "processed_data", "preprocessed_eye_data.Rdata"))
```
